\Title

### 2021-05-24 latex/paper/main.tex
Made edits up to (but not including) page 9 of paper_v2

The acrobot has inertia matrix 
\(M\), potential function \(V\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    M(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
For reasons that will become clear in later sections of this article, we will
use Hamiltonian mechanics to derive the dynamics of the acrobot. 
For this we require the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\).
The dynamics of the acrobot in \((q,p)\) coordinates are given in
\eqref{eqn:acrobot-hamiltonian}.
For shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau.
    \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.

@@

Our goal is to design a smooth function \(f : \Sone \times \R\) such that the
relation \(q_a = f(q_u,p_u)\) can be enforced asymptotically via feedback
control (in Section \ref{sec:vnhc} we call this a regular virtual
nonholonomic constraint), and the dynamics of the acrobot when the relation
holds gain energy on some set \(D\) in the sense of Definition
\ref{defn:energy-gain}.

@@

\subsection{Energy Regulation}
One can apply the results of Theorem \ref{thm:acrobot-energy-stabilization} 
towards energy regulation.

\textbf{TODO: move energy regulation theory from sims to here}
%For example, one might employ a supervisor which switches
%between injection and dissipation VNHCs to stabilize a desired energy
%level \(E_\text{des}\), with some hysteresis to avoid infinite switching.
%We will explore this idea in Section \ref{sec:simulations}.

@@

The dynamic parameters of Wang's acrobot can be found in Table
\ref{tab:acrobot-parameters}.
Evaluating the mechanical energy of an acrobot with these parameters
at the VNHC \(q_a = 0\) yields the mechanical energy of the nominal pendulum,
\[
    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
    .
\]

We will now simulate the effects of constraining Wang's acrobot with
\eqref{eqn:acrobot-constraint}, thereby demonstrating that VNHCs are robust to
model mismatch.
According to Theorem \ref{thm:acrobot-energy-stabilization}, the control
parameter \(I\) must be ``small" for our VNHC to inject (or dissipate) energy
into the simplified acrobot. 
Unfortunately, the Theorem does not specify how small \(|I|\) must be;
while we could make it arbitrarily small in simulations, we will eventually
implement this VNHC on a physical testbed where \(|I|\) must be large enough to
overcome friction.

@@

The level set \(E_\pi\) with energy \(E(\pi,0)\) is outlined in black.
Despite the change in acrobot model, this level set remains the boundary
between oscillations and rotations for the nominal pendulum.
The points where the orbit exits \(E_\pi\) are marked with black stars,
with the final departure marked by a red star.
Interestingly, our choice of \(I\) is large enough that we observe significant
differences between the nominal pendulum and the constrained dynamics:
\(E_\pi\) intersects the \(p_u\)-axis at \(\abs{p_u} \approx 0.18\), yet the
constrained acrobot begins rotating once it hits the
\(p_u\)-axis at \(\abs{p_u} \approx 0.16\). 
This indicates that higher values of \(I\) enable the acrobot to gain energy
faster and begin rotating sooner, so long as the actuator does not saturate.

@@

Once can use a supervisor to stabilize oscillations by appropriately
toggling between injection and dissipation VNHCs.
Specifically, one chooses a desired oscillation angle 
\(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a
hysteresis value \(\delta \in [0,1]\).
Each time the orbit crosses the \(q_u\)-axis (i.e. when \(p_u = 0\)) or 
when \(|q_u| = \pi\), the supervisor does the following: 
\begin{itemize}
    \item If \(|q_u| < (1-\delta)q_\text{des}\), enable the injection VNHC.
    \item If \(|q_u| > (1+\delta)q_\text{des}\), enable the dissipation VNHC.
    \item If \((1-\delta)q_\text{des} \leq |q_u| \leq (1+\delta)q_\text{des}\),
        keep the leg extended at \(q_a = 0\). This can be done continously since
        \(q_a = 0\) when \(p_u = 0\).
\end{itemize}

Figure \ref{fig:acrobot-osc-reg} shows the supervisor stabilizing an
oscillation with body angle \(q_\text{des} = \pi/2\) and a 
\(5\%\) hysteresis, meaning \(\delta = 0.05\).
The supervisor reevaluated its choice of VNHC at each black star;  
the red contour corresponds to the part of the orbit where the supervisor kept
the leg extended, because the oscillation was within tolerance of
\(q_\text{des}\).
The solid black line is the desired oscillation, and the dashed black
lines show the hysteresis around that orbit.

@@

\subsection{Rotation Regulation}
One can also use a supervisor to stabilize rotations.
In this case, one chooses a desired rotation rate \(p_\text{des} > 0\) and a
hysteresis value \(\delta \in [0,1]\).
Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
supervisor changes which VNHC is enforced as follows:
\begin{itemize}
    \item If \(|p_u| < (1-\delta)p_\text{des}\), enable the injection VNHC.
    \item If \(|p_u| > (1+\delta)p_\text{des}\), enable the dissipation VNHC.
    \item If \((1-\delta)p_\text{des} \leq |p_u| \leq (1+\delta)p_\text{des}\),
        extend the leg fully by setting \(q_a = 0\).
        For these simulations, we assume this can be done instantaneously.
\end{itemize}

Rotation regulation is demonstrated in Figure \ref{fig:acrobot-rot-reg}, where
the supervisor stabilizes \(p_\text{des} = 0.19\) with a \(2\%\) hysteresis 
\(\delta = 0.02\).
The supervisor changed its choice of VNHC at each black star until it was within
range of \(p_\text{des}\), at which point it extended the leg (shown in red).

>>>>

The acrobot has inertia matrix 
\(M\), potential function \(V\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    M(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
For reasons that will become clear in later sections of this article, we will
use Hamiltonian mechanics to derive the dynamics of the acrobot. 
For this we require the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\).
The dynamics of the acrobot in \((q,p)\) coordinates are given in
\eqref{eqn:acrobot-hamiltonian}.
For shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau.
    \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.
Let us now define what it means for a mechanical system to gain or lose energy.

@@

Our goal is to design a smooth function \(f : \Sone \times \R\) such that the
relation \(q_a = f(q_u,p_u)\) can be enforced asymptotically via feedback
control (in Section \ref{sec:vnhc} we call this a regular virtual
nonholonomic constraint).  
We will further require that the dynamics of the acrobot, when the relation
holds, gain or lose energy on some set \(D\) in the sense of Definition
\ref{defn:energy-gain}.

@@

\subsection{Energy Regulation}\label{sec:energy-reg}
One can apply the results of Theorem \ref{thm:acrobot-energy-stabilization} 
towards energy regulation;
that is, one can stabilize oscillations or rotations by appropriately toggling
between injection and dissipation VNHCs.

\textbf{Oscillation Regulation}: one chooses a desired oscillation angle 
\(q_\text{des} \in \, ]0,\pi[\) and, to avoid infinite switching, a
hysteresis value \(\delta \in [0,1]\).
Each time the orbit crosses the \(q_u\)-axis (i.e. when \(p_u = 0\)) or 
when \(|q_u| = \pi\), the supervisor does the following: 
\begin{itemize}
    \item If \(|q_u| < (1-\delta)q_\text{des}\), enable the injection VNHC.
    \item If \(|q_u| > (1+\delta)q_\text{des}\), enable the dissipation VNHC.
    \item If \((1-\delta)q_\text{des} \leq |q_u| \leq (1+\delta)q_\text{des}\),
        keep the leg extended at \(q_a = 0\). This can be done continously since
        \(q_a = 0\) when \(p_u = 0\).
\end{itemize}

\textbf{Rotation Regulation}: one chooses a desired rotation rate 
\(p_\text{des} > 0\) and a hysteresis value \(\delta \in [0,1]\).
Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
supervisor changes which VNHC is enforced as follows:
\begin{itemize}
    \item If \(|p_u| < (1-\delta)p_\text{des}\), enable the injection VNHC.
    \item If \(|p_u| > (1+\delta)p_\text{des}\), enable the dissipation VNHC.
    \item If \((1-\delta)p_\text{des} \leq |p_u| \leq (1+\delta)p_\text{des}\),
        extend the leg fully by setting \(q_a = 0\).
        For these simulations, we assume this can be done instantaneously.
\end{itemize}

@@

Recall that we defined the sets \(\mathcal{O}_1\) and
\(\mathcal{O}_2(\bar{\rho})\) used in Theorem \ref{thm:acrobot-energy-stabilization}
via the mechanical energy \eqref{eqn:acrobot-energy} of the nominal
pendulum associated with a simplified acrobot.
We obtained this nominal pendulum by setting \(I = 0\) in
\eqref{eqn:acrobot-constraint}.
Recall that the boundary between oscillations and rotations of the nominal
pendulum is obtained by finding the level set \(E_\pi\) of the nominal energy
corresponding to the upright equilibrium \((q_u,p_u) = (\pi,0)\).

We can extend this idea to find the nominal pendulum associated with Wang's
acrobot, whose physical parameters are provided in Table
\ref{tab:acrobot-parameters}.
Setting \(I = 0\) in \eqref{eqn:acrobot-constraint} for this acrobot
yields the following mechanical energy of the nominal pendulum:
\[
    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
    .
\]
Despite the change in acrobot model, the level set \(E_\pi\) with energy 
\(E(\pi,0)\) remains the boundary between oscillations and rotations
for the nominal pendulum.

We will now simulate the effects of constraining Wang's acrobot with the VNHC
\eqref{eqn:acrobot-constraint}, thereby demonstrating that VNHCs are robust to
model mismatch.
According to Theorem \ref{thm:acrobot-energy-stabilization}, the control
parameter \(I\) must be ``small" for our VNHC to inject (or dissipate) energy
into the simplified acrobot. 
Unfortunately, the Theorem does not specify how small \(|I|\) must be;
while we could make it arbitrarily small in simulations, we will eventually
implement this VNHC on a physical testbed where \(|I|\) must be large enough to
overcome friction.

@@

The level set \(E_\pi\) with energy \(E(\pi,0)\) is outlined in black.
Recall that this marks the boundary between oscillations and rotations of the
nominal pendulum.
The points where the orbit exits \(E_\pi\) are marked with black stars,
with the final departure marked by a red star.
Interestingly, our choice of \(I\) is large enough that we observe significant
differences between the nominal pendulum and the constrained dynamics:
\(E_\pi\) intersects the \(p_u\)-axis at \(\abs{p_u} \approx 0.18\), yet the
constrained acrobot begins rotating once it hits the
\(p_u\)-axis at \(\abs{p_u} \approx 0.16\). 
This indicates that higher values of \(I\) enable the acrobot to gain energy
faster and begin rotating sooner, so long as the actuator does not saturate.

@@

Recall from \ref{sec:energy-reg} that one can use a supervisor to stabilize
oscillations by appropriately toggling between injection and dissipation VNHCs.
Figure \ref{fig:acrobot-osc-reg} shows  stabilizing an
oscillation with body angle \(q_\text{des} = \pi/2\) and a 
\(5\%\) hysteresis, meaning \(\delta = 0.05\).
The supervisor reevaluated its choice of VNHC at each black star;  
the red contour corresponds to the part of the orbit where the supervisor kept
the leg extended, because the oscillation was within tolerance of
\(q_\text{des}\).
The solid black line is the desired oscillation, and the dashed black
lines show the hysteresis around that orbit.

@@

\subsection{Rotation Regulation}
One can also use a supervisor to stabilize rotations through the mechanism
described in Section \ref{sec:energy-reg}.
Rotation regulation for the acrobot is demonstrated in 
Figure \ref{fig:acrobot-rot-reg}, where the supervisor stabilizes 
\(p_\text{des} = 0.19\) with a \(2\%\) hysteresis \(\delta = 0.02\).
The supervisor changed its choice of VNHC at each black star until it was within
range of \(p_\text{des}\), at which point it extended the leg (shown in red).
